---
title: "Circle buffer and projections issue"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library("tidyverse")
library("sf")
library("mapview")
library("DT")
library("glue")
library("leaflet")
mapviewOptions(fgb = FALSE)
```

I'm struggling to figure out the most appropriate CRS to use for computing spheres of influence around arbitrary placed points on the globe. I haven't found a resource that recommends CRS for this purpose. I've made use of epsg.io which has a huge catalogue of CRS.

At the moment my workflow is experimental, I iterate through several CRS to find the best choice. Which is extremely far from efficient. This RMarkdown document shows my current workflow... the version I've posted to RPubs (https://rpubs.com/charliejhadley/circle-crs) has the code hidden for readability, the raw code can be found in this gist - https://gist.github.com/charliejhadley/d738166bee8719345d883d379d4f5587

# London

Let's start by storing the location of 221B Baker Street as a tibble with CRS 4236.

```{r echo=TRUE}
## Create an sf object storing the location of 221B Baker Street with EPSG:4326
baker_street <- tibble(lng = -0.158333, lat = 51.523333) %>% 
  st_as_sf(coords = c("lng", "lat"), crs = 4326)

baker_street %>% 
  mapview()
```

I've created `circle_buffer(point, crs_for_buffer, distance)` which will use `st_buffer()` to create a circle centered on `point` after reprojecting the data into `crs_for_buffer`

```{r echo=TRUE}
circle_buffer <- function(point,
                          crs_for_buffer,
                          distance,
                          crs_type) {
  crs_for_buffer <-
    ifelse(str_starts(crs_for_buffer, "[0-9]"),
           as.numeric(crs_for_buffer),
           crs_for_buffer)
  
  if (tolower(crs_type) == "geographic") {
    point %>%
      st_transform(crs_for_buffer) %>%
      st_buffer(dist = distance / 111320)
    
  } else {
    point %>%
      st_transform(crs_for_buffer) %>%
      st_buffer(dist = distance)
    
  }
  
}
```

We're going to be working with geographic and projected CRS which have different unit measures, degrees and metres respectively. Let's use 100e5 in projected CRS and 100e3/111320 in geographic CRS - as detailed in this SO answer [https://gis.stackexchange.com/a/142327/78822](https://gis.stackexchange.com/a/142327/78822).

Let's start off with CRS 4326, as expected we get highly eccentric ellipse instead of a circle:

```{r echo=TRUE}
circle_crs_4326 <- baker_street %>% 
  circle_buffer(4326, 100e3, "geographic")
circle_crs_4326 %>% 
  mapview()
```

I'm going to test the following CRS:

```{r, echo=FALSE}
london_projections <- tribble(
  ~crs_name, ~crs_link, ~crs_epsg, ~crs_for_code, ~proj_or_geo, ~details,
  "WGS 84", "https://epsg.io/102013", 4326, "4326", "Geographic", "",
  "British National Grid", "https://epsg.io/102013", 27700, "27700", "Projected", "",
  "Pseudo-Mercator -- Spherical Mercator", "https://epsg.io/3857", 3857, "3857", "Projected", "",
  "ED50 / UTM zone 30N", "https://epsg.io/23030", 23030, "23030", "Projected", "",
  "Europe Albers Equal Area Conic", "https://epsg.io/102013", NA, "+proj=aea +lat_1=43 +lat_2=62 +lat_0=30 +lon_0=10 +x_0=0 +y_0=0 +ellps=intl +units=m no_defs", "Projected", ""
) %>%
  mutate(projected_sf = pmap(list(crs_for_code, proj_or_geo), ~ circle_buffer(baker_street, .x, 100e3, .y)))

london_projections %>%
  select(-projected_sf) %>%
  mutate(crs_full_link = glue("<a href='{crs_link}'>{crs_name}</a>")) %>%
  select(crs_full_link, crs_epsg, proj_or_geo) %>%
  datatable(
    escape = FALSE,
    colnames = c("CRS Name", "EPSG Code", "Projected / Geographic"),
    rownames = FALSE,
    options = list(dom = "t")
  )

```

## Circle area **within** each CRS

The table below shows the computed area for each circle **within** each CRS:

```{r}
london_projections %>%
  mutate(area_within_crs = map_dbl(projected_sf, ~st_area(.x))) %>% 
  select(crs_name, area_within_crs) %>% 
  mutate(pi_deviance = scales::percent(1 - as.numeric(area_within_crs / 1e11 * pi), 0.01)) %>% 
  datatable(escape = FALSE, 
            colnames = c("CRS", "Area within CRS", "% variance from pi"),
            rownames = FALSE, 
            options = list(dom = "t"))
```

All four of the projected CRS deviate from Pi by the same amount - 1.3%

## Circle after reprojection to CRS 4326

Now we reproject each circle into CRS 4326 and recalculate area:

```{r}
london_projections %>%
  mutate(area_after_st_transform = map_dbl(c(projected_sf), ~st_area(st_transform(.x, 4326)))) %>% 
  select(crs_name, area_after_st_transform) %>% 
  mutate(pi_deviance = scales::percent(1 - as.numeric(area_after_st_transform / 1e11 * pi), 0.01)) %>% 
  arrange(desc(area_after_st_transform)) %>% 
  datatable(escape = FALSE, 
            colnames = c("CRS", "Area after st_transform()", "% variance from pi"),
            rownames = FALSE, 
            options = list(dom = "t"))
```

Visualising all of the circes it appears that it's a draw between: British National Grid, ED50 and the Europe Albers Equal Area Conic CRS.

```{r}
projected_circles_baker_street <- london_projections %>%
  mutate(area_after_st_transform = map_dbl(c(projected_sf), ~st_area(st_transform(.x, 4326)))) %>% 
  arrange(desc(area_after_st_transform)) %>% 
  mutate(transformed_4326_sf = map(c(projected_sf), ~st_transform(.x, 4326))) %>% 
  select(crs_name, transformed_4326_sf) %>% 
  unnest(cols = c(transformed_4326_sf)) %>% 
  st_bind_cols()

pal_circles_baker_street <- colorFactor("Set1", projected_circles_baker_street$crs_name)

projected_circles_baker_street %>% 
  leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>% 
  addPolygons(fillColor = ~pal_circles_baker_street(crs_name),
              fillOpacity = 1,
              weight = 1) %>% 
  addLegend(pal = pal_circles_baker_street,
            values = ~crs_name,
            opacity = 1)
```



# White House

Let's do exactly the same but with the White House.

```{r}
white_house <- tibble(lng = -77.0365, lat = 38.8977) %>% 
  st_as_sf(coords = c("lng", "lat"), crs = 4326)
```

Here are some CRS I chose:

```{r, echo=FALSE}
white_house_projections <- tribble(
  ~crs_name, ~crs_link, ~crs_epsg, ~crs_for_code, ~proj_or_geo, ~details,
  "WGS 84", "https://epsg.io/102013", 4326, "4326", "Geographic", "",
  "Pseudo-Mercator -- Spherical Mercator", "https://epsg.io/3857", 3857, "3857", "Projected", "",
  "US National Atlas Equal Area", "https://epsg.io/2163", 2163, "2163", "Projected", "",
  "North America Albers Equal Area Conic", "https://epsg.io/102008", NA, "+proj=aea +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80", "Projected", ""
) %>% 
  mutate(projected_sf = pmap(list(crs_for_code, proj_or_geo), ~circle_buffer(white_house, .x, 100e3, .y)))

white_house_projections %>% 
  select(-projected_sf) %>% 
  mutate(crs_full_link = glue("<a href='{crs_link}'>{crs_name}</a>")) %>% 
  select(crs_full_link, crs_epsg, proj_or_geo) %>% 
  datatable(escape = FALSE, 
            colnames = c("CRS Name", "EPSG Code", "Projected / Geographic"),
            rownames = FALSE, 
            options = list(dom = "t"))

```

## Circle area **within** each CRS

Once again we compute the area **within** each CRS:

```{r}
white_house_projections %>%
  mutate(area_within_crs = map_dbl(c(projected_sf), ~st_area(.x))) %>% 
  select(crs_name, area_within_crs) %>% 
  mutate(pi_deviance = scales::percent(1 - as.numeric(area_within_crs / 1e11 * pi), 0.01)) %>% 
  datatable(escape = FALSE, 
            colnames = c("CRS", "Area within CRS", "% variance from pi"),
            rownames = FALSE, 
            options = list(dom = "t"))
```

All four of the projected CRS deviate from Pi by the same amount - 1.35%

## Circle after reprojection to CRS 4326

Now we reproject each circle into CRS 4326 and recalculate area:

```{r}
white_house_projections %>%
  mutate(area_after_st_transform = map_dbl(c(projected_sf), ~st_area(st_transform(.x, 4326)))) %>% 
  select(crs_name, area_after_st_transform) %>% 
  mutate(pi_deviance = scales::percent(1 - as.numeric(area_after_st_transform / 1e11 * pi), 0.01)) %>% 
  arrange(desc(area_after_st_transform)) %>% 
  datatable(escape = FALSE, 
            colnames = c("CRS", "Area after st_transform()", "% variance from pi"),
            rownames = FALSE, 
            options = list(dom = "t"))
```

The least eccentric circle comes from the Pseudo-Mercator projection, all others are clearly eccentric.

```{r}
projected_circles_white_house <- white_house_projections %>%
  mutate(area_after_st_transform = map_dbl(c(projected_sf), ~st_area(st_transform(.x, 4326)))) %>% 
  arrange(desc(area_after_st_transform)) %>% 
  mutate(transformed_4326_sf = map(c(projected_sf), ~st_transform(.x, 4326))) %>% 
  select(crs_name, transformed_4326_sf) %>% 
  unnest(cols = c(transformed_4326_sf)) %>% 
  st_bind_cols()

pal_circles_white_house <- colorFactor("Set1", projected_circles_white_house$crs_name)

projected_circles_white_house %>% 
  leaflet() %>% 
  addProviderTiles(providers$OpenStreetMap) %>% 
  addPolygons(fillColor = ~pal_circles_white_house(crs_name),
              fillOpacity = 0.6,
              weight = 1) %>% 
  addLegend(pal = pal_circles_white_house,
            values = ~crs_name,
            opacity = 1)
```





















